---
title: "Introduction to R"
author: "ARF"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, comment = FALSE, warning = FALSE)
```


#Types of data
numeric

```{r}

is.numeric(2)

10-5

2+4

5-3

2*10

6/3

2.2/2

2^3

sqrt(4)

abs(-3)


```

# strings
```{r}

"Hello R"

is.character("Hello")

nchar("Hello") # in R we count from 1 :)
nchar("HelloR")
nchar("Hello R")

tolower("Hello R")
toupper("Hello R")

strtrim("Hello R",5)

"Hello R" + "Hello R"
paste("Hello", "R")
paste("Hello", "R", sep = "")
paste("Hello", "R", sep = ",")


```

# factors

```{r}

chocoBars <- factor(c("mars", "twix", "snickers", "milky way"))

is.factor(chocoBars)

table(chocoBars)

chocoBars <- factor(c("mars", "twix", "snickers", "milky way", "twix", "mars", "mars"))
str(chocoBars)
table(chocoBars)

chocoBars <- c("mars", "twix", "snickers", "milky way", "twix", "mars", "mars")
str(chocoBars)
table(chocoBars)

```


# logic

```{r}
1==0

is.logical(1==0)

1 | 0 # diff

1 > 0 # more than
1 < 0 # less than

2 <= 10
2 >= 10

# is.na() # is missing ?

# is.null() # is null ?

```


## Objects in R

# wektor
```{r}
# vector
c(1,3,5,7)

is.vector(c(1,3,5,7))

# sequences
seq(1, 100, by=1)


rep(1:3, times=3)

rep(1:2, each=3)

# length of vector
length(c(6,3,5,1))

# sort
sort(c(6,3,5,1),decreasing = TRUE)

sort(c("A", "C", "B"))

# what is inside vector
table(c(6,3,5,1,1,1,1,1))

# unique variables
unique(c(6,3,5,1,3,5,5))


```

# Variables assignemnt
```{r}
x <- c(1,3,4,6,10)
y <- c(1,3,4,9)

# select
x[4]
x[-4]
x[2:4]
x[-(2:4)]

x[c(1,5,2)]

x[x == 10]

x[x < 4]

t <- c("Tom", "Kate", "Peter", "Joe")

t[2:3]


x <- c(1,3,4,6,10)
y <- c(1,3,4,9,1)

min(x)
max(x)
sum(x)

x+y

z <- x+y
z

# combine
cbind(x,y)
rbind(x,y)

```


# data matrix

```{r}

X <- matrix(1:4, ncol = 2)
Y <- matrix(5:8, ncol=2)

Z <- matrix(1:28, nrow=4)
Z

t(X) # transpose

X+Y
Y/X
X*Y

X%*%Y # Matrix Multiplication

# solve(m, n)
# Find x in: m * x = n
solve(X, Y)

```

# List

```{r}

list <- list(name=c("Neil","Patrick"), surname="Harris", age=46,
inRelat=T)

list$name
list$name[1]

list[2]

l <- list(x = 1:5, y = c('a', 'b'))
l


l$x[2]

l[2]




```

# data frame !

```{r}

dt <- data.frame(id = c(1,2,3),
                 
                 age = c(25,21,22),
                 
                 is_boy = c(TRUE,TRUE,FALSE))
dt

class(dt)

dt[1:2, 1:3]

dt[ ,1:2]


```


## Tasks

1. Generate integers from 20 to 35 and save them in a vector called x.

2. Enter the y-vector below into R and calculate the sum of all elements, the difference between the largest and smallest y-values.
 y = c (1,8,2,6,3,8,5,5,5,5)

3. Generate a sequence of eight fours, then seven sixes and finally nine threes. Write it in an M matrix with six rows and four columns.

4. Create a data frame containing information about your closest family (name, age, gender, place of residence). Assign the created frame to the 'family' object.


## libraries

```{r}

install.packages("UsingR", dependencies = T)
 
library(UsingR)
 
?mean

help(package = "dplyr")

apropos("mean")

# Useful links
# https://www.rstudio.com/resources/cheatsheets/
# https://www.r-project.org/help.html

```



## Import your data

# from R libraries
```{r}
library(UsingR)

data(babyboom)

head(babyboom, 10) # first six rows
tail(babyboom) # last six rows


data("alltime.movies")
head(alltime.movies)

```

# csv file
```{r}
 read.csv(file,
         header=TRUE,
         sep=",",
         quote="\"",
         dec=".",
         fill=TRUE,
         comment.char="", ...)

```

# Excel file
```{r}

library(readxl)

# read_xls(path = ,sheet = , col_names = TRUE)

```

#TASK

6. Import file from repository


##### BREAK #####

install.packages("ggplot2", dependencies = T)
install.packages("dplyr", dependencies = T)
install.packages("tidyverse", dependencies = T)
install.packages("corrplot", dependencies = T)
install.packages("DataExplorer", dependencies = T)

