---
title: "Exploratory data analysis"
output: html_notebook
editor_options: 
  chunk_output_type: console
---

```{r}
library(ggplot2)
library(dplyr)
library(tidyverse)
library(corrplot)
library(DataExplorer)

```


```{r import data}

 read.csv("Indian Liver Patient Dataset.csv", header = TRUE, stringsAsFactors = TRUE, sep = ",")

data(mtcars)

```

```{r}
# Structure of your data

str(mtcars)

# missing values

is.na(mtcars$cyl)

colSums(sapply(mtcars, is.na))

mtcars <- na.omit(mtcars)

```

```{r column summary}

mean(mtcars$cyl)

max(mtcars$cyl)

summary(mtcars$cyl)
summary(mtcars)

describe(mtcars$cyl)
describe(mtcars)

table(mtcars$gear)

```

# Manipulation of data frame
```{r}

# index search
which(names(mtcars)=="cyl")
which(names(mtcars)=="mpg")
mtcars[, 1:6]

# better way :)

library(tidyverse)
mtcars <- mtcars %>% as_tibble()
select(mtcars, cyl:mpg)

# select
select(mtcars, cyl:carb) %>% filter(gear==3)

# select women over 80
select(mtcars, cyl:carb) %>% filter(vs==1 & gear > 2)

```



# Explore 

```{r boxplot}

boxplot(mtcars, col="green")

boxplot(mtcars$disp, col="blue")

```

```{r histogram}

hist(mtcars$qsec, col = "red")

```

```{r scatter plot}

plot(mtcars$cyl, mtcars$mpg)

```

# Correlation
```{r}

library(corrplot)

dt <- mtcars[c(2:5)]

c <- cor(dt, method = "pearson")

corrplot(c)


```

# 

```{r}

# table(mtcars$gender)
# 
# table(mtcars$is_patient)
# 
# barplot(table(mtcars$gender))
# barplot(table(mtcars$is_patient))

```


## nicer graphics

```{r}

plot(mtcars$cyl, mtcars$mpg)

 
plot(mtcars$cyl, mtcars$mpg,
     xlab="cyl",
     ylab="mpg")

plot(mtcars$cyl, mtcars$mpg,
     xlab="cyl",
     ylab="mpg",
     main="cyl vs. mpg level")

 
plot(mtcars$cyl, mtcars$mpg,
     xlab="cyl",
     ylab="mpg",
     main="cyl vs. mpg level",
     type = "p") #p, l, b, o, h, s, S

# symbol
plot(mtcars$cyl, mtcars$mpg,
    xlab="cyl",
     ylab="mpg",
     main="cyl vs. mpg level",
     type = "p",
     pch=15,
     col="red")

```

# Tasks

7. Create the mtcars6 frame from the mtcars frame, which only contains information about cars with 6 cylinders. Display table with basic statistics (for mtcars6 frame)

8. Create a graph showing the relationship between the two selected parameters.

9. Create a data frame containing the variables: age, tot_proteins, ag_ratio (only for men). Create a histogram for each of the variables, excluding men over 80 years of age.








```{r}

DataExplorer::plot_str(mtcars)

DataExplorer::plot_histogram(mtcars$disp)

DataExplorer::plot_correlation(mtcars)

DataExplorer::plot_missing(mtcars)

DataExplorer::plot_boxplot(mtcars, by = 5)

DataExplorer::create_report(mtcars)

```







